Blueprint for efficient nuclear spin characterization in color center

### 2023-08-22 article.tex
work on the text
@majidzahedian , please reply to XXX

In this work, we take the 23 nuclear spins cluster characterized in paper \cite{abobeih2019atomic} as a realistic case. 
First, we apply this to the dynamical decoupling sequence, and visualise the obtained results. 
Fig \ref{fig:3} compares DD signal for the same nuclear spin register but different defects in diamond. 
An intuitive description for the sensitivity of a sequence to the variation of hyperfine coupling could estimated knowing the width and sensitivity of the position of the resonances. 
First, we consider the case of NV center in diamond. 
The width of a dip $\delta \tau = \frac{A_{zx}}{\omega_L^2}$  and taking the derivative of the resonance time, we define the sensitivity of longitudinal hyperfine coupling as follows:
\begin{gather}
	\delta A_{zz} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zz}}}} = \frac{2}{\tau_k} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
\end{gather}
If we assume the typical values of $T_2^{HE} = 100 $ $\mu$s and $\omega_L=500 $ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz. On the other hand, for group IV detects in diamond, The resonance times is sensitive to the second order of $A_{zx}$ and no sensitivity to $A_{zz}$. Hence, even at large $\tau$ the nuclear spin not separated well. We define the sensitivity for the spin one half system as follows:
\begin{gather}
	\delta A_{zx} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
\end{gather}
Assuming the typical values of $T_2^{HE} = 100 $ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate.\\

We want to make a sequence that is limited by T1 instead of T2. The idea of 5-pulse ESEEM \cite{kasumaj20085} is that to create an entanglement (Hahn-echo) before the free evolution (Ramsey). In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix. Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure}[htbp]
	\begin{center}
		\includegraphics[width=0.9\columnwidth]{pict/eseem_compare2.pdf}
		\caption{Simulated FFT of ESEEM signal with a zoom into weakly coupled nuclear spins for the register reported \cite{abobeih2019atomic} assuming the electron spin is one-half. b. Enhancing the sensitivity to weakly coupled spins by increasing number of $\pi$-pulses to $N=72$ pulses in each entangling period. $\tau$ is set to the Larmor bright spot. Simulated Fisher information for the case of the electron spin c. one d. one-half system.}
		\label{fig:4}
	\end{center}
\end{figure}
Fig \ref{fig:4} shows the furrier transform of the signal. Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. Each nuclear spins have two peaks in the spectrum. Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. This fact makes their characterization rather straight forward. To observe the weakly coupled nuclei more pronouncedly, one can use DDESEEM sequence by adding more $\pi$ pulses in the entangling periods. To distinguish other nuclei, extra measurements is required. One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \cite{vorobyov2022addressing}. However, this method is rather time-consuming as two time variables has to be swept. We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\

>>>>

In this work, we take the 23 nuclear spins cluster characterized in paper \cite{abobeih2019atomic} as a realistic case. 
First, we apply this to the dynamical decoupling sequence, and visualise the obtained results. 
Fig \ref{fig:3} compares DD signal for the same nuclear spin register but different defects in diamond. 
An intuitive description for the sensitivity of a sequence to the variation of hyperfine coupling could estimated knowing the width and sensitivity of the position of the resonances. 
First, we consider the case of NV center in diamond. 
The width of a dip $\delta \tau = \frac{A_{zx}}{\omega_L^2}$  and taking the derivative of the resonance time, we define the sensitivity of longitudinal hyperfine coupling as follows:
\begin{gather}
	\delta A_{zz} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zz}}}} = \frac{2}{\tau_k} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
\end{gather}
If we assume the typical values of $T_2^{HE} = 100 $ $\mu$s and $\omega_L=500 $ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz. 
For $S=1/2$, e.g. for group IV defects in diamond, the resonance times is sensitive to the second order of $A_{zx}$ and no sensitivity to $A_{zz}$. 
Hence, even at large $\tau$ the resonances related to nuclear spins will not be separated well. 
We define the sensitivity for the spin one half system as follows:
\begin{gather}
	\delta A_{zx} = \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
\end{gather}
Assuming the typical values of $T_2^{HE} = 100 $ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate and $\approx 160$ times worse than for $S=1$.
To quantify the difference in sensitivity to estimating the hyperfine parameters, we perform calculation of the Fisher information matrix (FIM) for the 23 nuclear spin cluster, with 23 $A_{zz}$ and 23 $A_{zx}$ parameters (see Fig. \ref{fig:3}c,d). 
The fisher information matrix is estimated for the probability to measure state $|0\rangle$ and reads as: 
\begin{equation}
F_{ij}(\textbf{A})=\sum_\tau \frac{\partial p(0, \tau, \textbf{A})}{\partial A_i}\frac{\partial p(0, \tau, \textbf{A})}{ \partial A_j} \frac{1}{p (1-p)}.
\end{equation}
For the typical case of NV center, with $S=1$, since all the nuclear spin resonances are clearly resolved, the Fisher information matrix takes a diagonal shape, revealing a low covariance between the various nuclear spin resonances. 
The crosstalk is existing between $A_{zz}$ and $A_{zx}$ of same nuclear spin, which is because the position of the peaks depends on both values. 
The fisher information gives a bound for the precision of parameter estimation, known as Cramér–Rao bound: 
\begin{equation}
\delta A^2 \ge \frac{1}{F(A) N}
\end{equation}
For the presented protocol, the precision for various spins is limited by \vvv{XXX} for the S=1, for the N = 4000.

 The striking difference to the $S=1$ fisher information matrix, is the behaviour of the defect with $S=1/2$

On the other hand, a sequence that is limited by $T_1$ instead of $T_2$ is based on 3 and 5-pulse ESEEM. The main principle of 5-pulse ESEEM \cite{kasumaj20085} is to create an entanglement (via e.g. Hahn-echo block) before the free evolution of the nuclear spins and after make a second correlating sequence. 
In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix, thus limitied by $T_1$ time. 
Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. 
The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure}[htbp]
	\begin{center}
		\includegraphics[width=0.9\columnwidth]{pict/eseem_compare2.pdf}
		\caption{Simulated FFT of ESEEM signal with a zoom into weakly coupled nuclear spins for the register reported \cite{abobeih2019atomic} assuming the electron spin is one-half. b. Enhancing the sensitivity to weakly coupled spins by increasing number of $\pi$-pulses to $N=72$ pulses in each entangling period. $\tau$ is set to the Larmor bright spot. Simulated Fisher information for the case of the electron spin c. one d. one-half system.}
		\label{fig:4}
	\end{center}
\end{figure}
Fig \ref{fig:4} shows the Fourier transform of the signal. 
Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. 
Each nuclear spins have two peaks in the spectrum. 
Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. 
This fact makes their characterisation rather straight forward. 
To observe the weakly coupled nuclei more pronouncedly, one can use DD-ESEEM sequence by adding more $\pi$ pulses in the entangling periods. 
To distinguish other nuclei, extra measurements is required. 
One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \cite{vorobyov2022addressing}. 
However, this method is rather time-consuming as two time variables has to be swept. 
We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. 
Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\
